---
title: "Conference Registration"
output: 
  rmarkdown::html_document:
    highlight: pygments
    number_sections: no
    self_contained: yes
    toc: no
    toc_depth: 2
    toc_float: yes
---
```{r, echo=F, results=F, message=F, warning=F}
library(tibble)
library(dplyr)
library(readxl)
library(knitr)
library(kableExtra)

registration <- list()



registration$in_person <- tribble(
  ~"Type", ~"Regular", ~"IBS member", ~"Student",
  "Early-bird", 360, 260, 50,
  "Full", 440,	340,	75,
  "Course (per half-day)",  100, 75, 25
)

registration$virtual <- tribble(
  ~"Type", ~"Regular", ~"IBS member", ~"Student",
  "Early-bird", 180, 130, 25,
  "Full", 220,	170,	50,
  "Course (per half-day)",  75, 50, 10
)

registration$dinner <- tribble(
  ~"Type", ~"Regular", ~"IBS member", ~"Student",
  "Dinner (separate)", 60, 60, 30
) 

for (reg_type in names(registration)){
  registration[[reg_type]] <- registration[[reg_type]] %>%
    mutate(across(where(is.numeric), \(x){paste0(x, " CHF")}))
}

print_reg_table <- function(tbl) {
  kable(tbl, align="r") %>%
  column_spec(1, width = "40%") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
}
```
## Registration Desk
The registration desk is located in the center of the -1 (U1) level of Biozentrum, and will be open as follows:

* Sunday, 3<sup>rd</sup> of September: 08.00 - 18.00
* Montag, 4<sup>th</sup> of September: 07.00 - 18.00
* Tuesday, 5<sup>th</sup> of September: 08.00 - 18.00
* Wednesday, 6<sup>th</sup> of September: 08.00 - 14.30
* Thursday, 7<sup>th</sup> of September: 08.00 - 12.30

Note that no additional on-site tickets are available (see below).

## Tickets
Registration for the CEN2023 conference <b>[is handled by conftool](https://www.conftool.pro/cen2023/). Note that in-person tickets, the conference dinner, townhall reception and the go fastR course are sold out</b>. Regular registration fees now apply for virtual participation, see below for details. Discounted rates apply for students (up to and including during PhD, a file must be uploaded to confirm the student status) and IBS (International Biometric Society) members. 

If you are not an IBS member but wish to benefit from the reduced fees, then please apply first for IBS membership, e.g., following the processes outlined by [the IBS-ROeS](https://www.ibs-roes.org/mitgliedschaft/) or [the IBS-DR](http://www.biometrische-gesellschaft.de/en/die-gesellschaft/mitgliedschaft.html). After completing the membership registration, please register under the reduced membership fees below.

<br>

### In-person attendance
```{r, echo=F, include=T}
print_reg_table(registration$in_person) %>% row_spec(c(1,2,3), color = "lightgray")
```
### Virtual attendance
```{r, echo=F, include=T}
print_reg_table(registration$virtual) %>% row_spec(c(1, 3), color = "lightgray")
```

### Conference dinner
```{r, echo=F, include=T}
print_reg_table(registration$dinner) %>% row_spec(1, color = "lightgray")
```
<br>
